---
title: "Regression Models Project"
author: "Lucian Pop"
date: "Sunday, September 27, 2015"
output: pdf_document
---

## Executive Summary  
In this paper we will analyze mtcars data set and determine the relation between mpg and other significant variables. This data is extracted from the 1974 Motor Trend magazine, and denotes fuel consumption and other aspects of automobile performance for 32 73–74 car models.

Upon analysis was found that Manual transmission better Automatic transmission by about 7mpg given the same weight and power. 

## Load Data and Libraries
```{r, echo=FALSE}
library(ggplot2)
library(stats) 
data(mtcars)
```

## Exploratory Data Analysis  
```{r, echo=FALSE}
summary(cars)
```

At first we plotted mpg vs am (see Annex:Figure 1) and observed that overall automatic transmission achieve less mpg than manual transmission. However there might be other factors involved.
We further analyze mean mpg for "automatic"" vs "manual"" 
```{r }
means <- round(tapply(mtcars$mpg, mtcars$am, mean), 2)
```


We than proceed to pair plot variables  and observed that other factors influence mnps: hp, wt, disp, cyl.
Upon further analysis "cyl" can be discarded since "disp" is more important.
going further we will analyze models based on mpg, mpg+wt, mpg+wt+disp and mps+wt+hp.


```{r, echo=FALSE}
mtcars$cyl  <- factor(mtcars$cyl)
mtcars$vs   <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am   <- factor(mtcars$am,labels=c('Automatic','Manual'))
```


```{r, echo=FALSE}
means <- round(tapply(mtcars$mpg, mtcars$am, mean), 2)
```


```{r, echo=FALSE}
plot(cars)
```


```{r, echo=FALSE}
model_full       <- lm(mpg ~ ., mtcars)
model_am         <- lm(mpg ~ am , mtcars)
model_am_wt      <- lm(mpg ~ am + wt, mtcars)
model_am_wt_disp <- lm(mpg ~ am + wt + disp, mtcars)
model_am_wt_hp   <- lm(mpg ~ am + wt + hp, mtcars)

```


```{r, echo=FALSE}
summary(model_am)$coef
summary(model_am_wt)$coef
summary(model_am_wt_disp)$coef
summary(model_am_wt_hp)$coef
```



```{r, echo=FALSE}
anova(model_am, model_am_wt, model_am_wt_disp,model_am_wt_hp)

```


```{r, echo=FALSE}
confint(model_am_wt_hp)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



## Regression Analysis




## Residual Analysis and Diagnostics  

```{r, echo=FALSE}
sum((abs(dfbetas(model_am_wt_hp)))>1)
```

## Appendix: Figures  

1. Plot of MPG vs. Transmission  
```{r}
ggplot(mtcars, aes(y=mpg, x=factor(am, labels = c("automatic", "manual")), fill=(am)))+
  geom_violin(colour="red", size=1.5)+
  xlab("Transmission") + ylab("MPG")
```  
2. Pair Graph of Motor Trend Road Test Cars Data  
```{r}
pairs(mtcars, panel=panel.smooth, main='Road Test Cars Data', col=mtcars$am, pch=19)
```  
3. Scatter Plots MPG vs. factors  
```{r}
qplot(mpg, wt,  data=mtcars,colour=am )
qplot(mpg, disp,  data=mtcars,colour=am )
qplot(mpg, hp,  data=mtcars,colour=am )
```  
4. Residual Plots
```{r}
par(mfrow = c(2, 2))
plot(model_am_wt_hp)
```  